Adaptive Resolution Simulation: Theoretical Fundations and Open Questions

### 2014-10-15 writeDown/pub.epj/paper.tex
back up before thermodynamics

The potential interpolation approach defines the hybrid energy by the same idea:
\begin{align}\label{eqn:v-v-interpol}
  {V}_{\moleidxone \moleidxtwo}=w_\moleidxone w_\moleidxtwo{V}_{\moleidxone\moleidxtwo}^{\AT}+(1-w_\moleidxone w_\moleidxtwo){V}^{\CG}_{\moleidxone\moleidxtwo} 
\end{align}
The intermolecular force is therefore calculated by
\begin{align}\nonumber
  {\vect F}^V_{\moleidxone \moleidxtwo}&= -\nabla_{\moleidxone}{V}_{\moleidxone \moleidxtwo} \\ \label{eqn:f-v-interpol}
  &=w_\moleidxone w_\moleidxtwo{\vect F}_{\moleidxone\moleidxtwo}^{\AT}+(1-w_\moleidxone w_\moleidxtwo){\vect F}^{\CG}_{\moleidxone\moleidxtwo}  - \nabla w_\moleidxone\cdot w_\moleidxtwo (V^\AT_{\moleidxone \moleidxtwo} - V^\CG_{\moleidxone \moleidxtwo})
\end{align}
Where $\nabla w_\moleidxone$ is the gradient of the weighting function measured at position $\vect r_\moleidxone$,
and the superscipt ``$V$'' denotes that the force is defined
by the potential interpolation.
By  definition~\eqref{eqn:f-v-interpol}, the force of potential interpolation is conservative. 
The difference between the force and potential
interpolations lies in the last term  of Eq.~\eqref{eqn:f-v-interpol}, which is a force
acting along the direction of decreasing weighting function.
It should be noted that
the last term breaks the Newton's third law, therefore the force of potential interpolation
does not conserves the momentum.

@@

Eq.~\ref{eqn:dist-1-0} is further splitted as
\begin{align}
  p(\vect x_\AT \vert N_\AT) =
  \sum_{N_\HY} \int {d}\vect x_\HY\,
  p(\vect x_\AT \vert N_\AT; \vect x_\HY, N_\HY)\cdot
  p(\vect x_\HY, N_\HY\vert N_\AT)  
\end{align}
The first probability density in the integral is the probability density
of atomistic DOFs conditioned on the number of atomistic molecules and all
hybrid DOFs.
It can shown that if
(1) all interactions in the system are cut-offed;
(2) The atomistic region interacts with the bybrid region only in an atomistic way;
(3) The system is short-range correlated, and the correlation between the atomistic
and the coarse-grained regions is negligible, then the probability density $p(\vect x_\AT \vert N_\AT; \vect x_\HY, N_\HY)$
is approximated by
\begin{align}\label{eqn:dist-conf-0}
  p(\vect x_\AT \vert N_\AT; \vect x_\HY, N_\HY)
  \propto
  \exp\Big\{-\beta \mh^\AT(\vect x_\AT; \vect x_\HY, N_\HY) \Big\},
\end{align}
where 
\begin{align}
  \mh^\AT(\vect x_\AT; \vect x_\HY, N_\HY)
  =
  \sum_{i=1}^{N_\AT}\frac12 m_i\vect v_i^2 +
  \sum_{i,j=1}^{N_\AT} \frac12 V^\AT(\vect r_{ij}) +
  \sum_{i=1}^{N_\AT}\sum_{j=N_\AT+1}^{N_\AT + N_\HY} V^\AT(\vect r_{ij})
\end{align}
is the atomistic Hamiltonian with parameters $\vect x_\HY, N_\HY$. It
should be noted that the probability density~\eqref{eqn:dist-conf-0}
is identical to that of the reference system with a subregion that
corresponds to the hybrid region.

In general, the probability density $p(\vect x_\HY, N_\HY\vert N_\AT)$
is not the same as the reference system.  For example, it has been
shown that the density distribution in the hybrid region deviates from
the desired one, even if the coarse-grained side is modelled to
reporduce the atomistic pressure~\cite{poblete2010coupling}.  However,
it is possible to raise necessary conditions that systematically
improves the accuracy of the probability density. The first necessary
condition is that the density profile of the hybrid region should be a
constant that is identical to the density of the atomistic reference
at desired equilibrium:
\begin{align}
  \rho_\HY(\vect r) = \rho_\AT^\ast = \frac NV
\end{align}
The second necessary condition is that the two body probability
density in the hybrid region is the same as the full atomistic
reference, In homogeneous and isotropic system, it is equivalent to
ask that the hybrid radial distribution function (RDF) should be the
same as that of the reference system:
\begin{align}
  g_\HY(r) = g^\ast_\AT(r).
\end{align}
It is fully justified to systematically raise the necessary conditions
upto $m$-th multibody probability density, however, it is not convenient in
numerics, because the correction to the $m$-th multibody probability density
needs a $m$-body correction force in general, which is computationally
too expensive.

In practice, the density in the hybrid region is correct by the
thermodynamic force, which is applied on top of the AdResS intermolecular interactions:
\begin{align}
  \vect F_\moleidxone &= \sum_\moleidxtwo \vect F_{\moleidxone\moleidxtwo}  + \vect F_\moleidxone^\thf  \\
  \vect F^V_\moleidxone &= \sum_\moleidxtwo \vect F^V_{\moleidxone\moleidxtwo} + \vect F_\moleidxone^{\thf,V}
\end{align}
calculated by a iterative scheme:

Therefore
certain manipulation in the hybrid region is need improve the accuracy
of $p(\vect x_\HY, N_\HY\vert N_\AT)$. However, it is still not clear
what is the sufficient conditions to achieve good accuracy in $p(\vect x_\HY, N_\HY\vert N_\AT)$.
It is easy to raise necessary conditions: The marginal distributions
of the hybrid cooridnates
\begin{align}
  p(\vect x_\HY) = \sum_{N_\HY = 0}^\infty \sum_{N_\AT = 0}^\infty p(\vect x_\HY, N_\HY\vert N_\AT) p(N_\AT)
\end{align}
should be the same as the reference system. Here we use the fact that
the probability $ p(N_\AT)$ should be of good accuracy, which will be
discussed later. This approach is validated because the discussion of
accuracy of $ p(N_\AT)$ does not require a good accuracy of $ p(\vect
x_\HY, N_\HY\vert N_\AT)$.  The $m$th marginal distribution of a probability density saying $p(\vect x_\HY)$ is defined
by
\begin{align}
  p_\HY^{(m)}(\vect x_1, \cdots \vect x_m) = \int\cdots\int d\vect x_{m+1}\cdots d\vect x_{N_\HY}
  p (\vect x_1, \cdots, \vect x_m, \vect x_{m+1}, \cdots \vect x_{N_\HY})
\end{align}
In a homogeneous system, the first marginal distribution is the
density profile in the

@@

\section{Test}
citation~\cite{wang2014exploring}

>>>>

The potential interpolation approach defines the hybrid energy by the same idea:
\begin{align}\label{eqn:v-v-interpol}
  {V}_{\moleidxone \moleidxtwo}=w_\moleidxone w_\moleidxtwo{V}_{\moleidxone\moleidxtwo}^{\AT}+(1-w_\moleidxone w_\moleidxtwo){V}^{\CG}_{\moleidxone\moleidxtwo} 
\end{align}
The intermolecular force is therefore calculated by taking the negative gradient on the potential ${\vect F}^V_{\moleidxone \moleidxtwo}= -\nabla_{\moleidxone}{V}_{\moleidxone \moleidxtwo}$,
which is explicitly written as
\begin{align}\label{eqn:f-v-interpol}
  {\vect F}^V_{\moleidxone \moleidxtwo}
  &=w_\moleidxone w_\moleidxtwo{\vect F}_{\moleidxone\moleidxtwo}^{\AT}+(1-w_\moleidxone w_\moleidxtwo){\vect F}^{\CG}_{\moleidxone\moleidxtwo}  - \nabla w_\moleidxone\cdot w_\moleidxtwo (V^\AT_{\moleidxone \moleidxtwo} - V^\CG_{\moleidxone \moleidxtwo})
\end{align}
Where $\nabla w_\moleidxone$ is the gradient of the weighting function measured at position $\vect r_\moleidxone$,
and the superscipt ``$V$'' denotes that the force is defined
by the potential interpolation.
By  definition~\eqref{eqn:f-v-interpol}, the force of potential interpolation is conservative. 
The difference between the force and potential
interpolations lies in the last term  of Eq.~\eqref{eqn:f-v-interpol}, which is a force
acting along the direction of decreasing weighting function.
It should be noted that
the last term breaks the Newton's third law, therefore the force of potential interpolation
does not conserves the momentum.

@@

\subsection{The accuracy of the configurational probability density}
The configurational probability density~\ref{eqn:dist-1-0} is further splitted as
\begin{align}
  p(\vect x_\AT \vert N_\AT) =
  \sum_{N_\HY} \int {d}\vect x_\HY\,
  p(\vect x_\AT \vert N_\AT; \vect x_\HY, N_\HY)\cdot
  p(\vect x_\HY, N_\HY\vert N_\AT)  
\end{align}
The first probability density in the integral is the probability density
of atomistic DOFs conditioned on the number of atomistic molecules and all
hybrid DOFs.
It can shown that if
(1) all interactions in the system are cut-offed;
(2) The atomistic region interacts with the bybrid region only in an atomistic way;
(3) The system is short-range correlated, and the correlation between the atomistic
and the coarse-grained regions is negligible, then the probability density $p(\vect x_\AT \vert N_\AT; \vect x_\HY, N_\HY)$
is approximated by
\begin{align}\label{eqn:dist-conf-0}
  p(\vect x_\AT \vert N_\AT; \vect x_\HY, N_\HY)
  \propto
  \exp\Big\{-\beta \mh^\AT(\vect x_\AT; \vect x_\HY, N_\HY) \Big\},
\end{align}
where 
\begin{align}
  \mh^\AT(\vect x_\AT; \vect x_\HY, N_\HY)
  =
  \sum_{i=1}^{N_\AT}\frac12 m_i\vect v_i^2 +
  \sum_{i,j=1}^{N_\AT} \frac12 V^\AT(\vect r_{ij}) +
  \sum_{i=1}^{N_\AT}\sum_{j=N_\AT+1}^{N_\AT + N_\HY} V^\AT(\vect r_{ij})
\end{align}
is the atomistic Hamiltonian with parameters $\vect x_\HY, N_\HY$. It
should be noted that the probability density~\eqref{eqn:dist-conf-0}
is identical to that of the reference system with a subregion that
corresponds to the hybrid region.
The probability density and local Hamiltonian
can be written down for the coarse-grained region analogically.

In general, the probability density $p(\vect x_\HY, N_\HY\vert N_\AT)$
is not the same as the reference system.  For example, it has been
shown that the density distribution in the hybrid region deviates from
the desired one, even if the coarse-grained side is modelled to
reporduce the atomistic pressure~\cite{poblete2010coupling}.  However,
it is possible to raise necessary conditions that systematically
improves the accuracy of the probability density. The first necessary
condition is that the density profile of the hybrid region should be a
constant that is identical to the density of the atomistic reference
at desired equilibrium:
\begin{align}\label{eqn:necessary-1st}
  \rho_\HY(\vect r) = \rho_\AT^\ast = \frac NV
\end{align}
The second necessary condition is that the two body probability
density in the hybrid region is the same as the full atomistic
reference, In homogeneous and isotropic system, it is equivalent to
ask that the hybrid radial distribution function (RDF) should be the
same as that of the reference system:
\begin{align}\label{eqn:necessary-2nd}
  g_\HY(r) = g^\ast_\AT(r).
\end{align}
It is fully justified to systematically raise the necessary conditions
upto $m$-th multibody probability density, for example the third necessary condition would be
\begin{align}
  C^{(3)}_\HY = C^{(3)\ast}_\AT
\end{align}
where $C^{(3)}$ is the three-body correlation.
Since the system is homogeneous
and isotropic, the three-body probability density is equivalent to
three-body correlation function. 
% However, it is not convenient in
% numerics, because the correction to the $m$-th multibody probability density
% needs a $m$-body correction force in general, which is computationally
% too expensive.

In practice, the density in the hybrid region is correct by the
thermodynamic force~\cite{fritsch2012adaptive}, which is applied on top of the AdResS intermolecular interactions:
\begin{align}
  \vect F_\moleidxone &= \sum_\moleidxtwo \vect F_{\moleidxone\moleidxtwo}  + \vect F_\moleidxone^\thf,  \\
  \vect F^V_\moleidxone& = \sum_\moleidxtwo \vect F^V_{\moleidxone\moleidxtwo} + \vect F_\moleidxone^{\thf,V},
\end{align}
where the thermodynamic forces is a one-body force defined over space,
i.e.~$\vect F^\thf_i = \vect F^\thf(\vect r_\moleidxone)$ and $\vect F^{\thf, V}_\moleidxone = \vect F^{\thf,V}(\vect r_\moleidxone)$,
for force and potential interpolations, respectively.
The thermodynamic force is applied only in the hybrid region, and is
calculated by the following iterative scheme:
\begin{align}
  \vect F_{k+1}^{\thf(, V)} (\vect r) = \vect F_k^{\thf(, V)} (\vect r)-
  \frac{M}{\kappa(\rho^\ast_\AT)^2} \nabla\rho_k(\vect r)
\end{align}
where $k$ denotes the step of iteration. $\kappa$ is the
isothermo-compressibility. In equilibrium the density profile $\rho(\vect r)$ in the
atomistic and coarse-grained regions are constants, so the thermodynamic force
is updated only in the hybrid region. When the hybrid density is flat, the thermodynamic
force converges. An important subsequence of a flat hybrid density is that
\begin{align}
  \rho_\AT(\vect r) = \rho_\HY(\vect r) = \rho_\CG(\vect r) = \rho_\AT^\ast,
\end{align}
because the density profiles match at the atomistic-hybrid and
hybrid-coarse-grained boundaries.

In a numerical example of SPC/E water~\cite{berendsen1987missing}
system, it has been shown in Ref.~\cite{wang2013grand} that when the
thermodynamic force is applied, the AdResS atomistic region reporduces
mulit-body configurational probability density of the reference system
upto three-body correlation.
In principle, higher order of mulit-body correlation can be investigated in
a similar way. A surprising numerical observation is that although
only the thermodynamic force is applied, the second order necessary condition~\eqref{eqn:necessary-2nd}
is automatically fulfilled, because the RDF is identical to the reference system in the buffer hybrid region
$\{\vect r \,:\, \vect r \in\Omega_\HY, \dist(\vect r, \Omega_\AT) < r_c\}$ that is interacting with
the atomistic region. While the third necessary condition is not satisfied because deviation
in the three-body correlation presents in the buffer bybrid region.
It should be noted that there is no theoretical background why the second necessary
condition is satisfied by only using the thermodynamic force,
so one cannot expect the same benefit in other systems.

The second necessary condition~\eqref{eqn:necessary-2nd} is fulfilled by the
RDF correction $\vect F^\rdf$~\cite{wang2012adaptive}, which is a conservative two-body force added to the force interpolation~\eqref{eqn:f-f-interpol} as an extra term
that applies only in the hybrid region:
\begin{align}\label{eqn:rdf-corr-f}
  {\vect F}_{\moleidxone \moleidxtwo}
  =
  w_\moleidxone w_\moleidxtwo{\vect F}_{\moleidxone\moleidxtwo}^{\AT}
  +
  (1-w_\moleidxone w_\moleidxtwo){\vect F}^{\CG}_{\moleidxone\moleidxtwo}
  +
  w_\moleidxone w_\moleidxtwo(1-w_\moleidxone w_\moleidxtwo){\vect F}^{\rdf}_{\moleidxone\moleidxtwo}.
\end{align}
The RDF correction force is constructed from a RDF correction potential $V^\rdf$ by
\begin{align}
  \vect F_{\moleidxone\moleidxtwo}^\rdf = \vect F^\rdf(\vect r_{\moleidxone\moleidxtwo})
  = -\nabla_{\vect r} V^\rdf(r_{\moleidxone\moleidxtwo}),
\end{align}
which is calculated by the iterative Boltzmann inversion:
\begin{align}
  V_{k+1}^\rdf(r) = V_k^\rdf(r) + k_BT \ln \Big[\frac{g_k(r)}{g^\ast_\AT(r)}\Big] 
\end{align}
where $k$ is the step of iteration. When the RDF of $k$th iteration is
identical to that of the full atomistic reference $g^\ast_\AT(r)$, the
iteration converges. It has been shown that the iterative scheme of
thermodynamic force coupled with the iterative Boltzmann inversion RDF
is effective in correcting the density profile $g_\HY(\vect r)$ and
RDF $g_\HY(r)$ simultaneously~\cite{wang2012adaptive}.
For SPC/E water system when the RDF correction is applied, the third
necessary condition is automatically satisfied~\cite{wang2013grand}. Again, there is no
theoretical prove behind this phenomenum, and one cannot expect the same
benefit for other systems.

\noindent\textbf{Remark:} The RDF correction is developed only for the force interpolation scheme,
and a natural extension to the potential interpolation seems to be
\begin{align}
  {V}_{\moleidxone \moleidxtwo}
  =
  w_\moleidxone w_\moleidxtwo{V}_{\moleidxone\moleidxtwo}^{\AT}
  +
  (1-w_\moleidxone w_\moleidxtwo){V}^{\CG}_{\moleidxone\moleidxtwo}
  +
  w_\moleidxone w_\moleidxtwo(1-w_\moleidxone w_\moleidxtwo){V}^{\rdf}_{\moleidxone\moleidxtwo}.  
\end{align}
The effectiveness of this proposal has never been tested, and is still an open question.

\subsection{The accuracy of the number probability}

The accuracy of number probability $p(N_\AT)$ is investigated by a
Taylor expansion w.r.t.~the relative size of the atomistic region to
the whole system, which vanishes in the thermodynamic limit. The sufficient condition for the first order accuracy
asks for a balance of the chemical potential~\cite{wang2013grand}:
\begin{align}\label{eqn:mu-eq}
  \mu_\CG - \mu_\AT = \omega_0,
\end{align}
where $\omega_0$ corresponds to the work from the filter, on each
molecule that enters the atomistic region. An important conclusion
form Ref.~\cite{wang2013grand} is that the relation~\eqref{eqn:mu-eq}
holds when the thermodynamic force is applied so that the atomistic
and coarse-grained densities match. The conclusion is derived under
the assumption of thermodynamic limit and the decorrlation between the
atomistic and coarse-grained regions.

The second order accuracy is achieved if the isothermo-compressibilty of the atomistic and coarse-grained sizes matches:
\begin{align}
  \kappa_\AT = \kappa_\CG
\end{align}
This is achieved by coarse-grained modeling, for example, the
structure based coarse-grained models that reproduces the atomistic
RDF~\cite{wang2009comparative}.

\subsection{The WCA potential as a generic energy and particle reservior}

From the accuracy analysis one surprising fact is that if the
requirement for the accuracy is not very high, and only applying the
thermodynamic force is acceptable, then there is actually no
restriction on the coarse-grained model. One can even use ideal gas as
a coarse-grained model, but the computational difficulty lies in the
sudden switching on of the atomistic interaction when a coarse-grained
molecule enters the hybrid region. If two molecules are entering at
the same time and the same location, then the atomistic contribution
to the intermolecular interaction~\eqref{eqn:f-f-interpol} or
\eqref{eqn:f-v-interpol} is infinity, which drives the simulation
unstable. This numerical difficulty can be avoided by using capped
atomistic interaction~\cite{praprotnik2005adaptive}, or by using a
gradually switching-on core-softened atomistic
interaction~\cite{heyes2010thermodynamic}.
In Ref.~\cite{wang2013grand}, the authors instead tested with
Weeks-Chandler-Andersen (WCA) potential~\cite{weeks1971role} as the
coarse-grained model, which is a short-ranged and pure repulsive
interaction.  For a SPC/E water system, the cut-off radius of WCA potential
can be 2.4 time smaller than the cut-off used in the atomistic
region. Since the computational cost with is of order $\mo(r_c^3)$,
the computational cost spent on each pair interaction
is 19 times cheaper. For each pair of molecules the atomistic model computes
10 pairwise interactions (9 electrostatic + 1 van der Waals), while the
WCA model computes only one, therefore the WCA model in total costs $1/190$ 
computational effort on force computation than atomistic model.
The WCA approach has been successfully used in calculating the
chemical potential of various complex fluids and mixtures~\cite{agarwal2014chemical}.

% Therefore
% certain manipulation in the hybrid region is need improve the accuracy
% of $p(\vect x_\HY, N_\HY\vert N_\AT)$. However, it is still not clear
% what is the sufficient conditions to achieve good accuracy in $p(\vect x_\HY, N_\HY\vert N_\AT)$.
% It is easy to raise necessary conditions: The marginal distributions
% of the hybrid cooridnates
% \begin{align}
%   p(\vect x_\HY) = \sum_{N_\HY = 0}^\infty \sum_{N_\AT = 0}^\infty p(\vect x_\HY, N_\HY\vert N_\AT) p(N_\AT)
% \end{align}
% should be the same as the reference system. Here we use the fact that
% the probability $ p(N_\AT)$ should be of good accuracy, which will be
% discussed later. This approach is validated because the discussion of
% accuracy of $ p(N_\AT)$ does not require a good accuracy of $ p(\vect
% x_\HY, N_\HY\vert N_\AT)$.  The $m$th marginal distribution of a probability density saying $p(\vect x_\HY)$ is defined
% by
% \begin{align}
%   p_\HY^{(m)}(\vect x_1, \cdots \vect x_m) = \int\cdots\int d\vect x_{m+1}\cdots d\vect x_{N_\HY}
%   p (\vect x_1, \cdots, \vect x_m, \vect x_{m+1}, \cdots \vect x_{N_\HY})
% \end{align}
% In a homogeneous system, the first marginal distribution is the
% density profile in the

@@

For the 
In Ref.~\cite{wang2013grand}, it has been shown that

\section{Dynamical properties: adaptive resolution simulation beyond equilibrium}
citation~\cite{wang2014exploring}
